---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rdailychange

<!-- badges: start -->

<!-- badges: end -->

## Overview

The goal of rdailychange is to allow the user to mimic the shedding phenomenon 
of snow on solar panels by extracting sequential changes in snow water 
equivalent (SWE). Day-1 observations are sequential daily changes in SWE.
Day-2 observations assume that the shedding phenomenon occurs after two days. 
Hence, Day-2 observations are a sum of two consecutive days of positive 
sequential changes. Day-3, Day-4 and Day-5 observations are a sum of three,
four and five consecutive sequential daily changes. In addition, they allow 
for at most one middle sequential change to be a negative value as long as the 
sum of the other sequential daily changes are greater than the negative 
change. 



## Installation

### Development version

This package is still under development, to get a bug fix or to use a feature 
from the development version, you can install and load
the development version from [GitHub](https://github.com/) with:


``` r
# install.packages("devtools")
devtools::install_github("Kinekenneth48/rdailychange")
library(rdailychange)
```

## Package Contents

The package revolves around three main function named *extract_observations*,
*fit_observations*, and *plot_events*. Documentation about these functions can be 
obtained by entering the following in the R console

```{r example}
?rdailychange::extract_observations
?rdailychange::fit_observations
?rdailychange::plot

```

The package also contains a sample dataset of snow load measurements(SWE) 
of 9 weather stations from 1951 to 2020 snow year. The data 
is available by installing the rdailychange package and invoking the command.

```{r }
data("sample_data")
sample_data = force(sample_data)

dim(sample_data)
```


## Package Usage
Once installed, this package can extract snow load values that mimic 
shedding phenomenon of snow on solar panel and plot the mean recurrence interval
(MRI) per weather stations. The following example makes use of the *sample_data*
dataset. The example will following the workflow of extracting the Day-X method
observations, fitting those observations to either a generalized extreme 
value distribution or generalized Pareto distribution, and finally visualizing 
the MRI values from the weather stations. This example will focus on Day-1 method,
however, the same worflow can be applied to Day-2,-3,-4, and -5 methods. 


### Extract Day-1 method observations from the 7 weather stations
After extracting Day-1 method observations, a information about a 
typical weather stations will include the Day-1 observations and the 
annual maximum load for each snow year.

```{r}
d1 = rdailychange::extract_observations(sample_data, day=1)

#after
d1$USW00023062

```


### Fit the observations 
Once the observations are extracted, we can now fit a extreme value distribution
to the observations. The package can fit the Generalized Extreme Value(GEV) or
the Generalized Pareto distribution (GP), which are the two main distributions
for fitting extreme values. These distributions are preferred because of their
ability to better fit extreme values of observations. The return value will
include 50-year, 100-year, and 500-year MRI value from the Day-1 method along 
side their ratio with respective to the annual data.

```{r}
d1_parm = rdailychange::fit_observations(d1, type = "GEV")
d1_parm 
```


### Visualize MRI values
Finally, we can visualize the 50-year MRI values for each weather station.

```{r}
plot(d1_parm)
```




## Getting help

If you encounter a clear bug, please file an issue with a minimal reproducible 
example on [GitHub](https://github.com/Kinekenneth48/carsimr.pomeyie/issues).

---
Please note that this package is released under the
[MIT License](https://www.r-project.org/Licenses/MIT).
By participating or using this package you agree to abide by the terms of 
the license.
