---
title: "Introducting the Rdailychange package"
author: "Kenneth Kin Pomeyie, Braden Mindrum, Kinspride Duah"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Background

Snow loads have long been a primary design consideration for roofs, but little work has been done to determine the unique interactions of snow and solar panels. It is crucial that solar panels in many portions of the United States be built to withstand severe winter weather. A combination of tilting angles and the insulating property of solar panels ensures that snow is melted off panels within days. This shedding phenomenon limits the amount of time that snow accumulates on the panel, which is a critical structural difference compared to non-slip roofing systems. This shedding phenomenon of snow on panels motivates the need for a method to mimic and analyze snow measurements as though it felt on a solar panel. 
The <TT>rdailychange</TT> package provides a way to extract daily sequential changes, fit, and estimate ground snow load mean recurrence intervals (MRIs) for any weather station across the United States. The DAY-1 method assumes the shedding phenomenon of solar panels occurs daily, as such observations are daily sequential
changes. On the other hand, the DAY-1,DAY-2, DAY-3, DAY-4, and DAY-5 method assume snow shedding occurs after two  three , fourth, and fifth day of continuous snow accumulation. A simple demonstration of the workflow underlying estimating MRI values from single and multiple-day snow accumulations is provided in this vignette for the 50 FOS. The package README also demonstrates this workflow by estimating MRI values for 7 FOS weather stations. 



* * * 



## Data

The daily snow water equivalent of measurement stations used in this analysis was collected from the National Oceanic and Atmospheric Administration’s ([NOAA](https://www.ncdc.noaa.gov/cdo-web/search)) Global Historical Climatology Network using the snowload2 package. The data analysis will focus on snow water equivalent (SWE) for 251 National Weather Service First Order Stations (FOS) from October 1, 1950, to June 30, 2020, for a total of 70 snow seasons. Note that not all stations will have a total of 70 snow seasons. These data are included in the package and can be accessed with the commands


```{r}
library(rdailychange)
data("sample_data")
head(sample_data)
```

The following figure shows a map of 50 stations locations.The code for this map 
makes use of the ggplot2, sf, and dplyr package. 

```{r message=FALSE, warning=FALSE, fig.height = 6, fig.width = 6, fig.align = "center"}
library(ggplot2)
library(usmap)
library(dplyr)

df = distinct(sample_data, ID, .keep_all = TRUE) %>%
  select(LONGITUDE, LATITUDE)

 
transformed_data <- usmap_transform(df)

plot_usmap("states") + 
  geom_point(data = transformed_data, 
             aes(x = LONGITUDE.1, y = LATITUDE.1), 
             color = "red",
             size = 2)
```

* * * 

## Extract Day-X observations
Our goal is to mimic the snow shedding phenomenon on solar panel across the United States. Doing so 
allows us to compare how sensitive are MRI values from a single day to multiple day accumulation. 
The first step in the workflow is to extract the single and multiple-day observations.
Using the <TT>extract_observations()</TT> function as demonstrated below a 
list of list object is returned. Each list represents a weather station and within each list 
is the Day-X observations and the annual maximum snow loads. Note that as we move from
a single to  multiple day the number of extracted observations reduce. 



```{r}
# extract single-day observations
d1 = rdailychange::extract_observations(sample_data, day=1)
d1[[1]][[1]]


# extract Day-2 observations
d2 = rdailychange::extract_observations(sample_data, day=2)
d2[[1]][[1]]


# extract Day-3 observations
d3 = rdailychange::extract_observations(sample_data, day=3)
d3[[1]][[1]]
```

* * * 

## Fit Distributions
Once the observations are extracted, we can now fit a distribution
to the observations. The package can fit the Generalized Extreme Value(GEV) or
the Generalized Pareto distribution (GP), which are the two main distributions
for fitting extreme values. These distributions are preferred because of their
ability to better fit extreme values of observations. The return value will
include 50-year, 100-year, and 500-year MRI value from the Day-X method along 
with their ratio with respective to the annual data.

```{r}
d1_parm_gev = rdailychange::fit_observations(d1, type = "GEV")
d1_parm_gp = rdailychange::fit_observations(d1, type = "GP")
d1_parm_gev[1:2,]
d1_parm_gp[1:2,]
```

```{r}
d2_parm_gev = rdailychange::fit_observations(d2, type = "GEV")
d2_parm_gp = rdailychange::fit_observations(d2, type = "GP")
d2_parm_gev[1:2,]
d2_parm_gp[1:2,]
```

```{r}
d3_parm_gev = rdailychange::fit_observations(d3, type = "GEV")
d3_parm_gp = rdailychange::fit_observations(d3, type = "GP")
d3_parm_gev[1:2,]
d3_parm_gp[1:2,]
```

From the above output, we can see that fitting a GEV using a block-maxima approach or 
a GP using  the peak-over-thresholds approach did not significantly change our results. 

* * * 

## Visualize MRI values
In this section of the workflow we will visualize the 50-year MRI and their
respective ration of the solar-specific MRI values to that of the traditional design
load standard. The traditional design standard MRI values are calculated using annual
maximum loads. Note that the <TT>plot_events()</TT> function demonstrated below can be 
applied to multiple-day objects.

```{r fig.height = 6, fig.width = 6, fig.align = "center"}
plot_events(d1_parm_gev, event = 50, annual = FALSE)

```

```{r fig.height = 6, fig.width = 6, fig.align = "center"}
plot_events(d1_parm_gev, event = 50, annual = TRUE)

```



* * * 
